---
title: "Attitudes of Brazilian Voters on the United States and China"
subtitle: "How Partisanship and Presidential Approval Shape View of Foreign Policy"
author: "Gustavo Marangoni"
format: 
  beamer:
    theme: Madrid
    colortheme: crane
    aspectratio: 32
knitr:
  opts_chunk:
    echo: false
    eval: true
    prompt: false
    tidy: false
    comment: NA
    message: FALSE
    warning: FALSE
    fig.align: 'center'
---

```{r}
library(tidyverse)
library(readr)
library(dplyr)
library(knitr)
library(haven)

df <- readxl::read_xlsx("C:/Users/gugau/OneDrive/Área de Trabalho/Data Analysis/Problem Sets/PS4/F00018444-Latinobarometro_2023_Brazil_ExcelText_eng_v1.xlsx")

brazil <- df |> 
  select(
    `EDAD: Age`,
    `SEXO: Sex`,
    `REEEDUC.1: REEDUC 1`,
    `S2: People sometimes describe themselves as belonging to a social class. Would you describe yourself as belonging to`,
    `P15STGBS: Approval of the management of the government led by the President`,
    `P16ST: Self-positioning on Left-Right scale`,
    `P12ST: {COUNTRY} ruled by a few powerful groups for their own benefit or for the good of all the people`,
    `P23ST.A: Opinion about the United States`,
    `P23ST.C: Opinion about the China`,
    `P24ST.A: Rating of the relationship between the country and the United States.`,
    `P24ST.C: Rating of the relationship between the country and the China`,
    `P25N.A: Rating of the economic relations between the country and the United States`,
    `P25N.C: Rating of the economic relations between the country and China`,
    `P27SDN: {COUNTRY} should have a closer relationship with the United States or China`,
    `P43STGBS.A: If elections were this Sunday, which party would you vote for`,
    `P43STGBS.B: For coding only`,
  ) |> 
  rename(
    age            = `EDAD: Age`,
    sex            = `SEXO: Sex`,
    education      = `REEEDUC.1: REEDUC 1`,
    social_class   = `S2: People sometimes describe themselves as belonging to a social class. Would you describe yourself as belonging to`,
    pres_approval  = `P15STGBS: Approval of the management of the government led by the President`,
    ideology       = `P16ST: Self-positioning on Left-Right scale`,
    gov_perception = `P12ST: {COUNTRY} ruled by a few powerful groups for their own benefit or for the good of all the people`,
    favor_us       = `P23ST.A: Opinion about the United States`,
    favor_china    = `P23ST.C: Opinion about the China`,
    polrel_us      = `P24ST.A: Rating of the relationship between the country and the United States.`,
    polrel_china   = `P24ST.C: Rating of the relationship between the country and the China`,
    econrel_us     = `P25N.A: Rating of the economic relations between the country and the United States`,
    econrel_china  = `P25N.C: Rating of the economic relations between the country and China`,
    closer_to      = `P27SDN: {COUNTRY} should have a closer relationship with the United States or China`,
    party_vote     = `P43STGBS.A: If elections were this Sunday, which party would you vote for`,
    party          = `P43STGBS.B: For coding only`
  ) |> 
  filter(age >= 18) |> 
  mutate(
    favor_us1 = fct_relevel(
      favor_us, 
      "Very unfavorable", 
      "Somewhat unfavorable", 
      "Don't know",
      "Does not answer",
      "Somewhat favorable",
      "Very favorable"
    ),
    favor_china1 = fct_relevel(
      favor_china, 
      "Very unfavorable", 
      "Somewhat unfavorable", 
      "Somewhat favorable",
      "Very favorable"
    ),
    econrel_us1 = fct_relevel(
      econrel_us,
      "Very poor",
      "Fairly poor",
      "Fairly good",
      "Very good"
  ),
  econrel_china1 = fct_relevel(
      econrel_china,
      "Very poor",
      "Fairly poor",
      "Fairly good",
      "Very good"
  ),
  polrel_us1 = fct_relevel(
    polrel_us,
    "Very poor",
    "Fairly poor",
    "Fairly good",
    "Very good"
  ),
  polrel_china1 = fct_relevel(
    polrel_china,
    "Very poor",
    "Fairly poor",
    "Fairly good",
    "Very good"
  ),
   closer_to1 = fct_relevel(
     closer_to, 
     "Neither (if volunteered)",
     "Both (if volunteered)",
     "China",
     "United States"
   ),
  closer_to1 = fct_recode(
    closer_to1,
    "Neither" = "Neither (if volunteered)",
    "Both" = "Both (if volunteered)"
  ),
  party1 = case_when(
      party == "Voted government party"  ~ "Government",
      party == "Voted oppisition party" ~ "Opposition",
      TRUE ~ NA_character_
    )
  )

#Write a Function for Graphs for Categorical Variables (From PS3)
plot_us <- function(data, grouping, y_var){
  require(ggplot2)
  require(patchwork)
  require(dplyr)
  
  us_colors <- c("#002147", "#003F8A", "#0A5EB0", "#4A90D9", "#AEC6E8")
  
  plot <- 
    data |> 
    select({{ grouping }}, {{ y_var }}) |>
    filter(!{{ grouping }} %in% c("Don't know", "Does not answer", 
                                   "No answer", "DNK", "NA")) |>
    group_by({{ grouping }}) |> 
    mutate({{ grouping }} := as.factor({{ grouping }})) |> 
    ggplot(aes(x = {{ grouping }}, y = {{ y_var }})) +
    geom_col(aes(fill = {{ grouping }})) +
    scale_fill_manual(values = us_colors) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
    labs(caption = "Source: Latinobarómetro 2023") +
    theme_light() +
    theme(
      plot.title = element_text(
        size = 20,
        face = "bold",
        hjust = 0.5,
        color = "#002776"),
      plot.tag = element_text(
        face = "bold",
        size = 11,
        hjust = 1,
        vjust = 0,
        color = "#002776"),
      plot.tag.position = c(1, -0.05),
      plot.margin = margin(t = 10, r = 10, b = 100, l = 10),
      legend.position = "none",
      plot.subtitle = element_text(size = 12, color = "gray50"),
      panel.grid.major.y = element_line(color = "gray90"),
      panel.grid.major.x = element_blank(),
      axis.line = element_line(color = "#002776"),
      axis.text = element_text(color = "#002776", size = 11),
      axis.title = element_text(
        color = "#002776",
        face = "bold",
        size = 13),
      strip.background = element_rect(fill = "#002776"),
      strip.text = element_text(
        color = "white",
        face = "bold",
        size = 11)
    )
  
  return(plot)
}

plot_china <- function(data, grouping, y_var){
  require(ggplot2)
  require(patchwork)
  require(dplyr)
  
  china_colors <- c("#4B0000", "#8B0000", "#C0392B", "#E74C3C", "#F1948A")
  
  plot <- 
    data |> 
    select({{ grouping }}, {{ y_var }}) |>
    filter(!{{ grouping }} %in% c("Don't know", "Does not answer", 
                                   "No answer", "DNK", "NA")) |>
    group_by({{ grouping }}) |> 
    mutate({{ grouping }} := as.factor({{ grouping }})) |> 
    ggplot(aes(x = {{ grouping }}, y = {{ y_var }})) +
    geom_col(aes(fill = {{ grouping }})) +
    scale_fill_manual(values = china_colors) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
    labs(caption = "Source: Latinobarómetro 2023") +
    theme_light() +
    theme(
      plot.title = element_text(
        size = 20,
        face = "bold",
        hjust = 0.5,
        color = "#8B0000"),
      plot.tag = element_text(
        face = "bold",
        size = 11,
        hjust = 1,
        vjust = 0,
        color = "#8B0000"),
      plot.tag.position = c(1, -0.05),
      plot.margin = margin(t = 10, r = 10, b = 100, l = 10),
      legend.position = "none",
      plot.subtitle = element_text(size = 12, color = "gray50"),
      panel.grid.major.y = element_line(color = "gray90"),
      panel.grid.major.x = element_blank(),
      axis.line = element_line(color = "#8B0000"),
      axis.text = element_text(color = "#8B0000", size = 11),
      axis.title = element_text(
        color = "#8B0000",
        face = "bold",
        size = 13),
      strip.background = element_rect(fill = "#8B0000"),
      strip.text = element_text(
        color = "white",
        face = "bold",
        size = 11)
    )
  
  return(plot)
}
```

# **Overview**
- During Lula Inácio da Silva's administration, Brazil has deepened political and economic relationships with China, especially through the BRICS forum. 

- This is a contrast from Jair Bolsonaro's administration, who was openly allied to the United States and the Trump administration. 

- Given the upcoming **October presidential elections** and that both presidents represent different political parties and ideologies in Brazil, this study uses the Latinobarómetro to answer the question: 

### *To what extent does partisanship and presidential approval shape the view of the Brazilian population towards Foreign Policy, specifically in attitudes towards United States and China?*

# **Data and Methodology**
*Source: Latinobarómetro 2023 (During Lula's Administration)*

### Questions:
- Presidential approval
- U.S. and China favorability 
- Political and economic relations with U.S. and China 
- Great power preference 
- Vote intention / party alignment 

# **General Views Towards the United States**
```{r}
plot_us1 <- brazil |> 
  filter(!is.na(favor_us1)) |> 
  count(favor_us1) |> 
  plot_us(favor_us1, n) +
  labs(
    x = "Opinion of the United States",
    y = "Number of Respondents",
    title = "Brazilian Opinion of the United States"
    ) 

#| echo: true
plot_us1
```

# **General Views Towards China**
```{r}
plot_china1 <- brazil |> 
  filter(!is.na(favor_china1)) |> 
  count(favor_china1) |> 
  plot_china(favor_china1, n) +
   labs(
    x = "Opinion of China",
    y = "Number of Respondents",
    title = "Brazilian Opinion of China"
    )

#|echo: true
plot_china1
```

# **Economic and Political Relations Towards the U.S.**
```{r}
ploteconrel_us <- brazil |> 
  filter(!is.na(econrel_us1)) |> 
  count(econrel_us1) |> 
  plot_us(econrel_us1, n) +
  labs(
    x = "Economic Relationship",
    y = "Number of Respondents"
    )

polrel_usbrazil <- brazil |> 
  filter(!is.na(polrel_us1)) |> 
  count(polrel_us1) |> 
  plot_us(polrel_us1, n) +
  labs(
    x = "Political Relationship",
    y = "Number of Respondents"
    )

#|echo: true
ploteconrel_us + polrel_usbrazil +
  labs(
    title = "View on Economic and Political Relationship with the United States"
  ) +
  theme(
    plot.title = element_text(hjust = 1)
  )
```

# **Economic and Political Relations Towards China**
```{r}
ploteconrel_china <- brazil |> 
  filter(!is.na(econrel_china1)) |> 
  count(econrel_china1) |> 
  plot_china(econrel_china1, n) +
  labs(
    x = "Economic Relationship",
    y = "Number of Respondents"
    )

polrel_china <- brazil |> 
  filter(!is.na(polrel_china1)) |> 
  count(polrel_china1) |> 
  plot_china(polrel_china1, n) +
  labs(
    x = "Political Relationship",
    y = "Number of Respondents"
    )

#|echo: true
ploteconrel_china + polrel_china +
  labs(
    title = "View on Economic and Political Relationship with China"
  ) +
  theme(
    plot.title = element_text(hjust = 1.3)
  )
```

# **What are the Takeaways from These Graphs so far?**

Based on the plots analyzing favorability between the Great Powers, it seems that Brazilians tend to have a somewhat favorable view of the United States without accounting for partisanship. 

In contrast, when considering China, Brazilians tend to have a more unfavorable view than with the United States, but there is still some level of favorability towards the country as well. 

### In matters of political and economic relationships, there are two takeaways:
- Respondents tend to correlate political and economic relationships, making their answers fairly similar on both categories
- Most respondents have the impression that political and economic relationships with both countries are fairly good

> ***What happens when we account for partisanship?***

# **Partisanship and Foreign Policy Preferences in Brazil**
```{r}
#| fig.width: 14
#| fig.height: 8
plot_party_us <- brazil |>
  filter(!is.na(party1), !is.na(favor_us1)) |>
  filter(!favor_us1 %in% c("Don't know", "Does not answer")) |>
  count(party1, favor_us1) |>
  ggplot(aes(x = favor_us1, y = n, fill = party1)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c(
    "Government" = "#009C3B",
    "Opposition" = "#002776"
  )) +
  labs(
    title = "Opinion of the US",
    y = "Number of Respondents",
    fill = "Party"
  ) +
  theme_light() +
  theme(
      plot.title = element_text(
        size = 14,
        face = "bold",
        hjust = 0.5,
        color = "#002776"),
      plot.caption = element_text(size = 11),
      panel.grid.major.y = element_line(color = "gray90"),
      panel.grid.major.x = element_blank(),
      axis.line = element_line(color = "#002776"),
      axis.text = element_text(color = "#002776", size = 11),
      axis.title = element_text(
        color = "#002776",
        face = "bold",
        size = 13),
      strip.background = element_rect(fill = "#002776"),
      strip.text = element_text(
        color = "white",
        face = "bold",
        size = 11))

plot_party_china <- brazil |>
  filter(!is.na(party1), !is.na(favor_china1)) |>
  filter(!favor_china1 %in% c("Don't know", "Does not answer")) |> 
  count(party1, favor_china1) |>
  ggplot(aes(x = favor_china1, y = n, fill = party1)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c(
    "Government" = "#009C3B",
    "Opposition" = "#002776"
  )) +
  labs(
    title = "Opinion of China",
    y = "Number of Respondents",
    fill = "Party"
  ) +
  theme_light() +
  theme(
      plot.title = element_text(
        size = 14,
        face = "bold",
        hjust = 0.5,
        color = "#8B0000"),
      plot.tag = element_text(
        face = "bold",
        size = 11,
        hjust = 1,
        vjust = 0,
        color = "#8B0000"),
      panel.grid.major.y = element_line(color = "gray90"),
      panel.grid.major.x = element_blank(),
      axis.line = element_line(color = "#8B0000"),
      axis.text = element_text(color = "#8B0000", size = 11),
      axis.title = element_text(
        color = "#8B0000",
        face = "bold",
        size = 13),
      strip.background = element_rect(fill = "#8B0000"),
      strip.text = element_text(
        color = "white",
        face = "bold",
        size = 11))

plot_party_china + plot_party_us +
  plot_layout(guides = "collect") +
  plot_annotation(caption = "Source: Latinobarómetro 2023") &
  theme(plot.margin = margin(t = 10, r = 10, b = 30, l = 10),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        theme(
      plot.title = element_text(
        size = 14,
        face = "bold",
        hjust = 0.5,
        color = "#009C3B"),
      plot.caption = element_text(size = 11),
      panel.grid.major.y = element_line(color = "gray90"),
      panel.grid.major.x = element_blank(),
      axis.line = element_line(color = "#002776"),
      axis.text = element_text(color = "#002776", size = 11),
      axis.title = element_text(
        color = "#002776",
        face = "bold",
        size = 13),
      strip.background = element_rect(fill = "#009C3B"),
      strip.text = element_text(
        color = "white",
        face = "bold",
        size = 11)
    )
  )
```
# **Takeaways when Considering Partisanship**
## *What do Different Voters Think of the Great Powers?*
- Supporters of the Government Party (PT) have a slightly more favorable view of the United States than China
  - They also have a more unfavorable view of China than the opposition
- In proportion, government supporters have a similar level of support for both powers, but China has more respondents when accounting for an unfavorable opinion than the U.S.
- The opposition party clearly has a more positive view of the U.S, proportionally
  - In relation to China, there is some level of support, but there is a higher number of respondents with an unfavorable view of this great power

#  
### Proportionally, the highest difference is that the supporters of the government have a higher level of favorability to China than the opposition. 

**What happens when respondents are forced to choose a power for Brazil to align with?**

# **Should Brazil Get Closer to China or the US?**
```{r}
#| fig.width: 12
#| fig.height: 6
plot_closer <- brazil |>
  filter(!is.na(party1), !is.na(closer_to1)) |>
  filter(!closer_to %in% c("Don't know", "Does not answer")) |>
  count(party1, closer_to1) |>
  ggplot(aes(x = closer_to1, y = n, fill = party1)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c(
    "Government" = "#009C3B",
    "Opposition" = "#002776"
  )) +
  labs(
    title = "Preferred Great Power",
    y = "Number of Respondents",
    fill = "Party",
    caption = "Source: Latinobarómetro 2023"
  ) +
  theme_light() +
  theme(
      plot.title = element_text(
        size = 20,
        face = "bold",
        hjust = 0.5,
        color = "#009C3B"),
      plot.tag = element_text(
        face = "bold",
        size = 11,
        hjust = 1,
        vjust = 0,
        color = "#002776"),
      panel.grid.major.y = element_line(color = "gray90"),
      panel.grid.major.x = element_blank(),
      axis.line = element_line(color = "#002776"),
      axis.text = element_text(color = "#002776", size = 11),
      axis.title = element_text(
        color = "#002776",
        face = "bold",
        size = 13),
      strip.background = element_rect(fill = "#009C3B"),
      strip.text = element_text(
        color = "white",
        face = "bold",
        size = 11),
      axis.title.x = element_blank(),
      plot.margin = margin(t = 10, r = 10, b = 10, l = 10))

plot_closer
```
# **Who do Brazilians Choose to Align with?**
## **Align with the United States!**
- When accounting for partisanship, both sets of respondents proportionally favored the United States before China
  - Government supporters had more respondents favoring China, but still a majority choosing the United States over China

# **Conclusion**
- The overview indicated that Brazilians generally have favorable views of the United States and China, but more unfavorable views of China than the United States
- Considering partisanship, voters generally preferred an alignment with the United States before China, if they had to choose, but there were some differences among parties:

### Answer to initial Research Question:
> **Partisanship is correlated with Brazilian's views on foreign policy attitudes. Opposition voters consistently favored alignment with the U.S, while government supporters showed relatively more openness to China**

# **What to Expect of Brazil in the Future?**
### How is this relevant now?
- Given the importance of foreign policy, the upcoming presidential elections in **October 2026** will be crucial for determining Brazil's future internationally.
- **Lula's reelection** could determine a deepening alliance with China and a gradual shift of the balance of power in China's favor
- A victory from the opposition could result in a deepening alliance with the United States, which would maintain the U.S's hegemony in the Western Hemisphere
- Either outcome will have serious consequences not only for Brazil, but also to the broader **geopolitical competition** between the U.S and China in Latin America
